Assessing primitives performance\\ on multi-stage execution

### 2018-05-13 paper.tex
paper.tex

The Smart Syntax Plugin is one of the version existing until we started this work. The idea behind Smart Syntax was the following: the programmer can write code in Smalltalk using exclusively constructs that are recognized by the Smart Syntax compiler and that can be directly compiled to C (computing the size of a byte array, accessing its fields and doing integer arithmetic for example). With the addition of type annotations in the form of pragmas \cite{Duca16a} \footnote{Pragma are a form of method annotation.}, the Smart Syntax plugin was able to generate C code from the Smalltalk code. With this approach, the programmer has to write the optional primitive code a single time in Smalltalk. This code is used both as the fall-back code if the VM does not implement the primitive and at Slang to C compilation time to generate a C plugin from the code and the type annotations. This primitive follows the old specification, the full code is available in Appendix \ref{SmartSyntaxAppendix}.

\paragraph{Existing: Slang Plugin.}

The Slang plugin is a recent complete rewrite of the Smart Syntax plugin. The main difference is that the Smalltalk version is used only as fall-back code, while at Slang to C compilation time the VM uses code written in Slang to generate the C code. One of the side-effect is that the plugin's code is now part of the VM code base and not the client code bases, which is convenient for maintenance since the Cog VM has multiple client languages. However, the implementor now needs to maintain and evolve two version of the primitive, the fall-back code in Smalltalk and the code in Slang. This primitive follows the old specification, the full code is available in Appendix \ref{PluginAppendix}.

>>>>

The Smart Syntax Plugin is one of the version \sd{of what} existing until we started this work. The idea behind Smart Syntax is the following: the programmer writes code in Smalltalk using exclusively constructs that are recognized by the Smart Syntax compiler and that can be directly compiled to C (\emph{e.g.}, computing the size of a byte array, accessing its fields and doing integer arithmetic). With the addition of type annotations in the form of pragmas \cite{Duca16a} \footnote{Pragma are method annotation.}, the Smart Syntax plugin is able to generate C code from the Smalltalk code. \sd{A word on QuickTalk \cite{Ball86a}}
With this approach, the programmer has to write the optional primitive code a single time in Smalltalk. This code is used both as the fall-back code if the VM does not implement the primitive and at Slang to C compilation time to generate a C plugin from the code and the type annotations. This primitive follows the old specification, the full code is available in Appendix~\ref{SmartSyntaxAppendix}.
\sd{do we have a ref to the quickTalk approach? Because it was like that too if I remember.}

\paragraph{Existing: Slang Plugin.}
The Slang plugin is a recent complete rewrite of the Smart Syntax plugin. The main difference is that the Smalltalk version is used only as fall-back code, while at Slang-to-C compilation time the VM uses code written in Slang to generate the C code. One of the side-effect is that the plugin's code is now part of the VM code base and not the client code bases, which is convenient for maintenance since the Cog VM has multiple client languages. However, the implementor now needs to maintain and evolve two version of the primitive, the fall-back code in Smalltalk and the code in Slang. \sd{do we need the next sentence?}This primitive follows the old specification, the full code is available in Appendix~\ref{PluginAppendix}.
